---
title: "Characterization of the PRs"
author: "omitted"
date: "2024-04-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(lubridate)
library(GGally)

theme_set(theme_bw())
prs_characterization <- read.csv("../data/prs_characterization.csv")

prs_characterization <- prs_characterization %>%
    mutate(
        code_churn = additions + deletions,
        merged_at = as_datetime(merged_at),
        created_at = as_datetime(created_at),
        duration = difftime(created_at, merged_at, units = "days")
    )

prs_characterization <- prs_characterization %>%
    mutate(
        duration = abs(as.numeric(duration)),
    )
```

# Data overview

```{r}
glimpse(prs_characterization)
```

## PRs count

```{r}
prs_characterization %>% count()
```

## PRs count by repo

```{r}
prs_characterization %>%
    group_by(repo) %>%
    count()
```

## Commits count

```{r}
sum(prs_characterization$commits_count)
```

## Commits count by repo

```{r}
prs_characterization %>%
    group_by(repo) %>%
    summarise(sum(commits_count))
```

## General data statistics by repo

```{r}
prs_characterization %>%
    group_by(repo) %>%
    summarise(
        min_date = min(merged_at),
        max_date = max(merged_at),
        age_in_days = difftime(min(merged_at), max(merged_at), units = "days")
    ) %>%
    mutate(
    age_in_days = abs(as.numeric(age_in_days)),
    age_in_years = abs(as.numeric(age_in_days))/365.25
    )

prs_characterization %>%
    group_by(repo) %>%
    summarise(
        mean_commits = mean(commits_count),
        median_commits = median(commits_count),
        mean_changed_files = mean(changed_files_count),
        median_changed_files = median(changed_files_count)
    )

prs_characterization %>%
    group_by(repo) %>%
    summarise(
        mean_code_churn = mean(code_churn),
        median_code_churn = median(code_churn),
        mean_duration = mean(duration)
    )
```

## NCLOC and classes count by repo

* All values are related to the last commit of the last PR that was successfully executed by SonarQube.

```{r}
data.frame(repo = c("accumulo", "cayenne", "commons-collections","commons-io", "commons-lang", "helix", "httpcomponents-client", "maven-surefire", "opennlp", "struts", "wicket", "zookeeper"), NCLOC = c(440441, 318428, 67690, 30501, 95929, 189487, 76964, 110481, 155900, 234331, 251505, 131712), classes = c(5164, 4716, 839, 288, 918, 2106, 879, 3036, 2478, 3419, 5254, 1542))
```

## PRs and commits count through the years

```{r}
prs_characterization %>%
    mutate(year = lubridate::year(merged_at)) %>%
    group_by(year, repo) %>%
    summarise(PRs_quantity = n(), .groups = 'drop') %>%
    ggplot(aes(x = year, y = PRs_quantity, color=repo)) + 
    geom_line() +
    labs(x = "Year", y = "PRs count") + 
    ggtitle("PRs count by year") +
    scale_x_continuous(breaks = unique(lubridate::year(prs_characterization$merged_at)))

prs_characterization %>%
    mutate(year = lubridate::year(merged_at)) %>%
    group_by(year, repo) %>%
    summarise(commits_quantity = sum(commits_count), .groups = 'drop') %>%
    ggplot(aes(x = year, y = commits_quantity, color=repo)) + 
    geom_line() +
    labs(x = "Year", y = "Commits count") + 
    ggtitle("Commits count by year") +
    scale_x_continuous(breaks = unique(lubridate::year(prs_characterization$merged_at)))
```

# Distribution of features

## Commits count by PR

```{r}
tapply(prs_characterization$commits_count, prs_characterization$repo, summary)

prs_characterization %>%
    ggplot(aes(x = "", y = commits_count, color=repo)) +
      geom_boxplot() +
      facet_wrap(~repo) +
      labs(title = "Commits count by PR ~ repo")
```

## Code churn by PR

```{r}
tapply(prs_characterization$code_churn, prs_characterization$repo, summary)

prs_characterization %>%
    ggplot(aes(x = "", y = code_churn, color=repo)) +
      geom_boxplot() +
      facet_wrap(~repo) +
      labs(title = "Code churn by PR ~ repo")
```

## Added lines

```{r}
tapply(prs_characterization$additions, prs_characterization$repo, summary)

prs_characterization %>%
    ggplot(aes(x = "", y = additions, color=repo)) +
      geom_boxplot() +
      facet_wrap(~repo) +
      labs(title = "Added lines by PR ~ repo")
```

## Deleted lines

```{r}
tapply(prs_characterization$deletions, prs_characterization$repo, summary)

prs_characterization %>%
    ggplot(aes(x = "", y = deletions, color=repo)) +
      geom_boxplot() +
      facet_wrap(~repo) +
      labs(title = "Deleted lines by PR ~ repo")
```

## Changed files count

```{r}
tapply(prs_characterization$changed_files_count, prs_characterization$repo, summary)

prs_characterization %>%
    ggplot(aes(y = changed_files_count, color=repo)) +
      geom_boxplot() +
      facet_wrap(~repo) +
      labs(title = "Changed files count by PR ~ repo")
```

## PR duration

```{r}
tapply(prs_characterization$duration, prs_characterization$repo, summary)

prs_characterization %>%
    ggplot(aes(y = duration, color=repo)) +
      geom_boxplot() +
      facet_wrap(~repo) +
      labs(title = "PR duration ~ repo")
```

# Scatters

```{r}
prs_characterization %>%
    ggplot(aes(x = duration, y = commits_count, color=repo)) + 
    geom_point(alpha=0.75) +
    labs(x = "Duration", y = "Commits count") + 
    ggtitle("Commits count per duration")

prs_characterization %>%
    ggplot(aes(x = duration, y = changed_files_count, color=repo)) + 
    geom_point(alpha=0.75) +
    labs(x = "Duration", y = "Changed files count") + 
    ggtitle("Changed files count per duration")

prs_characterization %>%
    ggplot(aes(x = duration, y = code_churn, color=repo)) + 
    geom_point(alpha=0.75) +
    labs(x = "Duration", y = "Code churn") + 
    ggtitle("Code churn per duration")

prs_characterization %>%
    ggplot(aes(x = changed_files_count, y = code_churn, color=repo)) + 
    geom_point(alpha=0.75) +
    labs(x = "Changed files count", y = "Code churn") + 
    ggtitle("Code churn per changed files count")
```

# Correlations

```{r}
prs_characterization %>%
    select(additions, deletions, changed_files_count, commits_count, code_churn, duration) %>%
    ggpairs(title="Correlation between features")
```
