---
title: "Monitoring Analysis"
author: "omitted"
date: "2024-04-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(scales)
library(tidyverse)

theme_set(theme_bw())

monitoring_data <- read.csv("../data/prs_monitoring.csv")
monitoring_data$duration_formatted <- as.POSIXct(strptime(monitoring_data$duration, format = "%H:%M:%OS"))

duration_to_seconds <- function(duration_string) {
  parts <- strsplit(duration_string, "[:,.]")[[1]]
  hours <- as.numeric(parts[1])
  minutes <- as.numeric(parts[2])
  seconds <- as.numeric(parts[3])
  milliseconds <- as.numeric(parts[4])
  return(hours * 3600 + minutes * 60 + seconds + milliseconds / 1000)
}

monitoring_data$duration_in_seconds <- sapply(monitoring_data$duration, duration_to_seconds)
```

# Total analyzed commits
```{r}
nrow(monitoring_data)
```

# Total time to execute all commits analysis
```{r}
total_in_seconds <- sum(monitoring_data$duration_in_seconds)
total_days <- total_in_seconds / (60*60*24)
total_days
```

# Execution duration of commits

```{r}
monitoring_data %>%
    select(duration_formatted) %>%
    summary()
```

# Execution duration of commits by repo

```{r}
monitoring_data %>%
    group_by(repo) %>%
    pivot_wider(names_from = repo, values_from = c("duration_formatted")) %>%
    select("accumulo", "cayenne", "commons-collections", "commons-io", "commons-lang", "helix", "httpcomponents-client", "maven-surefire", "opennlp", "struts", "wicket", "zookeeper") %>%
    summary()
```

## Mean duration of commits by PR - Accumulo

PRs with pr_number >= 2000 and pr_number < 3000 was executed on Intel Core i5-12500H cores, while the rest on Ryzen 5 5600G cores.


```{r}
monitoring_data %>%
    filter(repo == "accumulo") %>%
    group_by(pr_number) %>%
    summarise(mean_duration = mean(duration_formatted)) %>%
    ggplot(aes(y=mean_duration, x=pr_number)) +
    geom_point() +
     labs(title = "Mean duration of commits by PR - Accumulo",
       x = "PR Number",
       y = "Mean duration")
```

## Mean duration of commits by PR - Helix

PRs with pr_number <= 2549 was executed on Intel Core i5-12500H cores, while the rest on Ryzen 5 5600G cores.

```{r}
monitoring_data %>%
    filter(repo == "helix") %>%
    group_by(pr_number) %>%
    summarise(mean_duration = mean(duration_formatted)) %>%
    ggplot(aes(y=mean_duration, x=pr_number)) +
    geom_point() +
     labs(title = "Mean duration of commits by PR - Helix",
       x = "PR Number",
       y = "Mean duration")
```