---
title: "RQ Answers"
author: "omitted"
date: "2024-04-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pct_format = scales::percent_format(accuracy = .01)

library(knitr)
library(dplyr)
library(ggplot2)
library(scales)
library(tidyverse)
library(ggsci)

ggplot2::ggplot()

options(kableExtra.latex.next.table.width = "0.5\\linewidth")
options(dplyr.summarise.inform = FALSE)

theme_set(theme_bw())

all_data <- read.csv("../data/processed_issues.csv", encoding = "UTF-8")
```


# Data overview

```{r}
glimpse(all_data)
```

# RQ1

## TDD (Technical Debt Density) by PR Distribution
```{r, echo = FALSE}
all_data %>%
    group_by(repo, pr_number) %>%
    summarise(total_debt = sum(debt), .groups = 'drop') %>%
    left_join(all_data %>%
              group_by(repo, pr_number) %>%
              select(file, ncloc_affected_file, repo, pr_number) %>%
              distinct(file, .keep_all = TRUE) %>%
              summarise(total_ncloc = sum(ncloc_affected_file), .groups = 'drop'),
            by = c("repo", "pr_number")) %>%
    ungroup() %>%
    mutate(tdd = total_debt / total_ncloc) %>%
    ggplot(aes(x=tdd, after_stat(density))) +
    geom_density(fill="steelblue", alpha = .75) +
    geom_vline(aes(xintercept=mean(tdd), color="Mean"), alpha = .75, linetype="dashed", linewidth=1) +
    geom_vline(aes(xintercept=median(tdd), color="Median"), alpha = .75,linetype="dashed", linewidth=1) +
    geom_text(aes(x=0, y= 0, label=paste("Mean:", round(mean(tdd), 4))), vjust=-28, hjust = -4.8, color="black") +
    geom_text(aes(x=0, y= 0, label=paste("Median:", round(median(tdd), 4))), vjust=-26, hjust = -4.2, color="black") +
    scale_color_manual(name = "Statistics", values = c(Median = "springgreen2", Mean = "red")) +
    labs(
        #title = "TDD per PR Distribution",
        x = "TDD"
    )
```


## TDV (Technical Debt Variation)

### TDV Distribution

```{r, echo = FALSE}
all_data %>%
    group_by(pr_number, repo) %>%
    summarise(
        new_unfixed_debt = sum(debt[status == "OPEN" & origin == "NEW"]),
        preexisting_fixed_debt = sum(debt[status == "CLOSED" & origin == "PRE-EXISTING"]),
        tdv = new_unfixed_debt - preexisting_fixed_debt,
        .groups = "drop"
    ) %>%
    select(pr_number, repo, tdv) %>%
    ungroup() %>%
    ggplot(aes(x = tdv, after_stat(density))) +
    geom_density(fill = "steelblue", alpha = .75) +
    geom_vline(aes(xintercept=median(tdv), color="Median"), alpha = .75,linetype="dashed", linewidth=1) +
    geom_text(aes(x=0, y= 0, label=paste("Mean:", round(mean(tdv), 4))), vjust=-28, hjust = -0.8, color="black") +
    geom_text(aes(x=0, y= 0, label=paste("Median:", round(median(tdv), 4))), vjust=-26, hjust = -1.6, color="black") +
    scale_color_manual(name = "Statistics", values = c(Median = "red")) +
    labs(
        #title = "TDV per PR Distribution",
        x = "TDV"
    )
```

### Percentages for TDV < 0 (TD decrease), TDV = 0 (TD unchanged) and TDV > 0 (TD increased) by repo
```{r, echo = FALSE}
all_data %>%
    group_by(pr_number, repo) %>%
    summarise(
        new_unfixed_debt = sum(debt[status == "OPEN" & origin == "NEW"]),
        preexisting_fixed_debt = sum(debt[status == "CLOSED" & origin == "PRE-EXISTING"]),
        tdv = new_unfixed_debt - preexisting_fixed_debt,
        .groups = "drop"
    ) %>%
    group_by(repo) %>%
    summarise(
        tdv_lower_than_zero = round((sum(tdv < 0) / n()) * 100, 2),
        tdv_equal_to_zero = round((sum(tdv == 0) / n()) * 100, 2),
        tdv_greater_than_zero = round((sum(tdv > 0) / n()) * 100, 2)
    )
```

### Mean of percentages of TDV (mean of repo percentages)
```{r echo=FALSE}
all_data %>%
    group_by(pr_number, repo) %>%
    summarise(
        new_unfixed_debt = sum(debt[status == "OPEN" & origin == "NEW"]),
        preexisting_fixed_debt = sum(debt[status == "CLOSED" & origin == "PRE-EXISTING"]),
        tdv = new_unfixed_debt - preexisting_fixed_debt,
        .groups = "drop"
    ) %>%
    group_by(repo) %>%
    summarise(
        tdv_lower_than_zero = round((sum(tdv < 0) / n()) * 100, 2),
        tdv_equal_to_zero = round((sum(tdv == 0) / n()) * 100, 2),
        tdv_greater_than_zero = round((sum(tdv > 0) / n()) * 100, 2)
    ) %>%
    summarise(
        mean_tdv_lower_than_zero = round(mean(tdv_lower_than_zero), 2),
        mean_tdv_equal_to_zero = round(mean(tdv_equal_to_zero), 2),
        mean_tdv_greater_than_zero = round(mean(tdv_greater_than_zero), 2)
    )
```

### Percentages of pre-existing TDV by repo

```{r echo=FALSE}
all_data %>%
    filter(origin == "PRE-EXISTING") %>%
    group_by(pr_number, repo) %>%
    summarise(
        preexisting_fixed_debt = sum(debt[status == "CLOSED" & origin == "PRE-EXISTING"]),
        preexisting_tdv = preexisting_fixed_debt,
        .groups = "drop"
    ) %>%
    group_by(repo) %>%
    summarize(
        preexisting_tdv_equal_to_zero = round((sum(preexisting_tdv == 0) / n()) * 100, 2),
        preexisting_tdv_greater_than_zero = round((sum(preexisting_tdv > 0) / n()) * 100, 2)
    )
```

## Mean of percentages of pre-existing TDV (mean of repo percentages)
```{r echo=FALSE}
all_data %>%
    filter(origin == "PRE-EXISTING") %>%
    group_by(pr_number, repo) %>%
    summarise(
        preexisting_fixed_debt = sum(debt[status == "CLOSED" & origin == "PRE-EXISTING"]),
        preexisting_tdv = preexisting_fixed_debt,
        .groups = "drop"
    ) %>%
    group_by(repo) %>%
    summarize(
        preexisting_tdv_equal_to_zero = (sum(preexisting_tdv == 0) / n()) * 100,
        preexisting_tdv_greater_than_zero = (sum(preexisting_tdv > 0) / n()) * 100
    ) %>%
    summarise(
        preexisting_tdv_equal_to_zero = round(mean(preexisting_tdv_equal_to_zero), 2),
        mean_preexisting_tdv_greater_than_zero = round(mean(preexisting_tdv_greater_than_zero), 2),
    )
```

### Percentages for new TDV by repo

```{r echo=FALSE}
all_data %>%
    filter(origin == "NEW") %>%
    group_by(pr_number, repo) %>%
    summarise(
        new_fixed_debt = sum(debt[status == "CLOSED" & origin == "NEW"]),
        new_tdv = new_fixed_debt,
        .groups = "drop"
    ) %>%
    group_by(repo) %>%
    summarize(
        new_tdv_equal_to_zero = round((sum(new_tdv == 0) / n()) * 100, 2),
        new_tdv_greater_than_zero = round((sum(new_tdv > 0) / n()) * 100, 2)
    ) %>%
    group_by(repo)
```

## Mean of percentages of new TDV (mean of repo percentages)
```{r echo=FALSE}
all_data %>%
    filter(origin == "NEW") %>%
    group_by(pr_number, repo) %>%
    summarise(
        new_fixed_debt = sum(debt[status == "CLOSED" & origin == "NEW"]),
        new_tdv = new_fixed_debt,
        .groups = "drop"
    ) %>%
    group_by(repo) %>%
    summarize(
        new_tdv_equal_to_zero = round((sum(new_tdv == 0) / n()) * 100, 2),
        new_tdv_greater_than_zero = round((sum(new_tdv > 0) / n()) * 100, 2)
    ) %>%
    summarise(
        mean_new_tdv_equal_to_zero = round(mean(new_tdv_equal_to_zero), 2),
        mean_new_tdv_greater_than_zero = round(mean(new_tdv_greater_than_zero), 2),
    )
```

## RQ2
As data is unbalanced across repositories, as shown in the issue characterization (../issues_characterization.Rmd), we will examine the top 10 fixed issues and the top 10 unfixed issues for each repository, and then aggregate them into a rank using the $PositionScore$ metric given by the following formula:

$PositionScore_{i} = \sum\limits_{j=1}^{k} Position_{j}$

given a rule $i$, its $PositionScore_{i}$ will be the sum of its position across all $k$ repositiries.After calculating the metrics for all rules, we rank them in a TOP 10 of the most frequent rules (fixed or unfixed, depending on the context) across repositories. The lower the PositionScore, the more frequent the rule across repositories.

### TOP 10 fixed per repo ranked by count

All top 10 fixed rules for each repo.

```{r, echo = FALSE}
all_data %>%
    filter(status == "CLOSED") %>%
    group_by(repo, rule) %>%
    summarise(count = n()) %>%
    mutate(rank = rank(-count, ties.method = "random")) %>%
    filter(rank <= 10) %>%
    group_by(repo, rank) %>%
    mutate(rule_count = paste(rule, " - ", count)) %>%
    select(repo, rule_count, rank) %>%
    pivot_wider(names_from = repo, values_from = rule_count) %>%
    arrange(rank)
```


### TOP 10 fixed

TOP 10 fixed rules by PositionScore metric.

```{r, echo = FALSE}
positions_fixed <- data.frame(
  rule = character(),
  position = integer(),
  stringsAsFactors = FALSE
)

for (current_repo in unique(all_data$repo)) {
    mapped_fixed_rules <- all_data %>%
        filter(repo == current_repo & status == "CLOSED") %>%
        group_by(rule, repo) %>%
        summarise(count = n()) %>%
        arrange(desc(count)) %>%
        head(10) %>%
        ungroup() %>%
        mutate(position = row_number()) %>%
        select(rule, position)
    
    positions_fixed <- bind_rows(positions_fixed, mapped_fixed_rules)
}

freq <- table(positions_fixed$rule)

missing_rules <- names(freq)[freq < 12]

missing_data <- data.frame(rule = character(), position = integer())

for (rule in missing_rules) {
  missing_count <- 12 - freq[rule]
  missing_instances <- data.frame(rule = rep(rule, missing_count), position = rep(11, missing_count))
  missing_data <- rbind(missing_data, missing_instances)
}

positions_fixed <- rbind(positions_fixed, missing_data)

positions_fixed %>%
  group_by(rule) %>%
  summarise(
     position_score = sum(position),
  ) %>% arrange(position_score)
```


### TOP 10 unfixed per repo ranked by count

All top 10 unfixed rules for each repo.


```{r, echo = FALSE}
all_data %>%
    filter(status == "OPEN") %>%
    group_by(repo, rule) %>%
    summarise(count = n()) %>%
    mutate(rank = rank(-count, ties.method = "random")) %>%
    filter(rank <= 10) %>%
    group_by(repo, rank) %>%
    mutate(rule_count = paste(rule, " - ", count)) %>%
    select(repo, rule_count, rank) %>%
    pivot_wider(names_from = repo, values_from = rule_count) %>%
    arrange(rank)

```

### TOP 10 unfixed

TOP 10 unfixed rules by PositionScore metric.

```{r, echo = FALSE}
positions_unfixed <- data.frame(
  rule = character(),
  position = integer(),
  stringsAsFactors = FALSE
)

for (current_repo in unique(all_data$repo)) {
    mapped_unfixed_rules <- all_data %>%
        filter(repo == current_repo & status == "OPEN") %>%
        group_by(rule, repo) %>%
        summarise(count = n()) %>%
        arrange(desc(count)) %>%
        head(10) %>%
        ungroup() %>%
        mutate(position = row_number()) %>%
        select(rule, position)
    
    positions_unfixed <- bind_rows(positions_unfixed, mapped_unfixed_rules)
}

freq <- table(positions_unfixed$rule)

missing_rules <- names(freq)[freq < 12]

missing_data <- data.frame(rule = character(), position = integer())

for (rule in missing_rules) {
  missing_count <- 12 - freq[rule]
  missing_instances <- data.frame(rule = rep(rule, missing_count), position = rep(11, missing_count))
  missing_data <- rbind(missing_data, missing_instances)
}

positions_unfixed <- rbind(positions_unfixed, missing_data)

positions_unfixed %>%
  group_by(rule) %>%
  summarise(
     position_score = sum(position),
  ) %>% arrange(position_score)
```


# Outliers

## Higher TDD

```{r, echo = FALSE}
all_data %>%
    group_by(repo, pr_number) %>%
    summarise(total_debt = sum(debt), .groups = 'drop') %>%
    left_join(all_data %>%
              group_by(repo, pr_number) %>%
              select(file, ncloc_affected_file, repo, pr_number) %>%
              distinct(file, .keep_all = TRUE) %>%
              summarise(total_ncloc = sum(ncloc_affected_file), .groups = 'drop'),
            by = c("repo", "pr_number")) %>%
    ungroup() %>%
    mutate(tdd = total_debt / total_ncloc) %>% arrange(desc(tdd))
```


## Higher TDV
```{r, echo = FALSE}
all_data %>%
    group_by(pr_number, repo) %>%
    summarise(
        new_unfixed_debt = sum(debt[status == "OPEN" & origin == "NEW"]),
        preexisting_fixed_debt = sum(debt[status == "CLOSED" & origin == "PRE-EXISTING"]),
        tdv = new_unfixed_debt - preexisting_fixed_debt,
        .groups = "drop"
    ) %>%
    select(pr_number, repo, tdv) %>%
    arrange(desc(tdv))
```


## Lower TDV
```{r, echo = FALSE}
all_data %>%
    group_by(pr_number, repo) %>%
    summarise(
        new_unfixed_debt = sum(debt[status == "OPEN" & origin == "NEW"]),
        preexisting_fixed_debt = sum(debt[status == "CLOSED" & origin == "PRE-EXISTING"]),
        tdv = new_unfixed_debt - preexisting_fixed_debt,
        .groups = "drop"
    ) %>%
    select(pr_number, repo, tdv) %>%
    arrange(tdv)
```

# Extra

## Number of PRs with pre-existing TD
```{r, echo = FALSE}
all_data %>% filter(origin == "PRE-EXISTING") %>% distinct(repo,pr_number) %>% count()
```
## Number of java:S1192 issues that affect test code

```{r, echo = FALSE}
all_data %>% filter(rule == "java:S1192") %>% filter(grepl("test", file)) %>% count()
```

## Number of java:S1192 issues that affect not test code
```{r, echo = FALSE}
all_data %>% filter(rule == "java:S1192") %>% filter(!grepl("test", file)) %>% count()
```

## Summary NCLOC
```{r, echo = FALSE}
all_data %>%
    group_by(pr_number, repo) %>%
    summarise(
        ncloc = sum(unique(ncloc_affected_file) %>% unlist()),
        debt = sum(debt),
        tdd = debt/ncloc,
        .groups = "drop"
    ) %>% select(ncloc) %>%
    summary()
```
